### 2022-04-26 paper/table-data.tex
final polishing

\begin{tabular}{ccrrrr}
\toprule                                                                        
automata set & level & marks & reduction [\%] & time [s] & \makecell{solver \\[-1ex] timeout} \\

\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2dela}\\129 automata\\523 acc.~marks}}
&
level 1 &
387~ & 26.0~~~~~~ & 755.4~ &  18~~~
\\
&
level 2 &
390~ & 25.4~~~~~~ & 1024.5~ &  20~~~
\\
&
level 3 &
393~ & 24.9~~~~~~ & 7273.3~ &  25~~~
\\
&
multi &
371~ & 29.1~~~~~~ & 9111.4~ &  21~~~
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2dgra}\\234 automata\\882 acc.~marks}}
&
level 1 &
552~ & 37.4~~~~~~ & 738.7~ & 14~~~
\\
&
level 2 &
569~ & 35.5~~~~~~ & 907.3~ &  18~~~
\\
&
level 3 &
561~ & 36.4~~~~~~ & 1163.4~ &  21~~~
\\
&
multi &
542~ & 38.5~~~~~~ & 2320.7~ &  21~~~
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl2tgba}\\70 automata\\198 acc.~marks}}
&
level 1 &
198~ & 0.0~~~~~~ & 49.3~ &  0~~~
\\
&
level 2 &
198~ & 0.0~~~~~~ & 78.9~ & 0~~~
\\
&
level 3 &
189~ & 4.5~~~~~~ & 1222.8~ & 7~~~
\\
&
multi &
189~ & 4.5~~~~~~ & 1259.1~ & 6~~~
\\
\midrule
\multirow{4}{*}{\makecell{\texttt{ltl3tela}\\91 automata\\348 acc.~marks}}
&
level 1 &
332~ & 4.6~~~~~~ & 465.9~ &  12~~~
\\
&
level 2 &
334~ & 4.0~~~~~~ & 553.9~ &  15~~~
\\
&
level 3 &
326~ & 6.3~~~~~~ & 691.8~ &  18~~~
\\
&
multi &
323~ & 7.2~~~~~~ & 1545.9~ & 18~~~
\\

\bottomrule                                                                     
\end{tabular}